---
title: "WUP_04"
author: "Marcia Sanchez"
date: "Summer 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
require(stats)
thedata<-read.csv(file="GSScsv.csv")
```





# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

First we wanted to find out what sore of data we had. We lean on writeout number 3 and I read the data and change the charaacter variables into factor variables and saw we had a numerical data which was wordsum and we now have a factor viaraibel called degreee. so we deducted that we have a numerical twitle categorical   
  
  
# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
ggplot(thedata, aes(x = DEGREE, y = WORDSUM)) +
geom_boxplot(fill = "skyblue", color = "darkblue", outlier.shape = NA) +
theme_minimal() +
labs(x = "DEGREE", y = "WORDSUM") +
theme(plot.title = element_text(hjust = 0.5)) +
ggtitle("Fancy Box Plot")
```
I see 5 comparisons among Bachelor, Graduate, High School, Junior College, LT High School. Bachelor and Junior college  are at the same lines levels. LT High School is the lowest of them all. The second comparison is High school is right after LT High School.

### Numerical Descriptive Results

```{r}
favstats(WORDSUM~DEGREE,data=thedata)
```

The average score on the graduates is a little over 10. There is a sure difference between high school and graduates. You get clarification when you run via anova. Statistically you can see that Graduate and LT High School is a big difference.

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

## Inferential Results

```{r}
model1 <- aov(WORDSUM~DEGREE,data=thedata)
summary.aov(model1)
```

```{r}

# Perform one-way ANOVA
model <- aov(WORDSUM ~ DEGREE, data = thedata)

# Conduct post-hoc test (Tukey's HSD)
posthoc <- TukeyHSD(model)

# Print the results
print(posthoc)
```

```{r}
# Assuming you have the "thedata" data frame with "DEGREE" and "WORDSUM" columns

# Load the emmeans package
library(emmeans)

# Perform one-way ANOVA
result <- aov(WORDSUM ~ DEGREE, data = thedata)

# Perform Scheffe's post hoc test
comparison <- emmeans(result, pairwise ~ DEGREE, adjust = "scheffe")

# Print the comparison table
print(comparison)
```


<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

There is a significance difference in statistical results between the schooling of all 5 degrees. The Hypothesis is null because the probability is very slim at 0.000236. High school is different from Graduate school. All the rest is 0.5 in P value. There are only 2 that are different Graduate and LT High School.

# Discussion and Conclusion

It is easy to look at Tukey HSD because Scheffe's.Test was a bit harder. I learned to ask chat GpT for another multiple comparisons. We also utilized our previous write up #3 to get data and formulas to work on the Write up #4. We tested many different possibilities and came to a conclusion that Tukey HSD is easier. I am learning to read data and create a simple way to interpret. In general of all the 5 grades  the most difference is between graduate and high school and graduate and LT High School. 

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
